\startYear{2023 May}

\subsection*{Section A -- Multiple-choice Questions}
\begin{multicols*}{2}

\begin{qsolution}{2023}{May}{Q1}{\tgElectronics \tgElectronicsSemiconductor}
    \mcqAns{C}

    \begin{enumualpha}
        \item Bandgap energy in semiconductors are non-zero in order to control the flow of current, so A is incorrect.
        \item Apart from group IV elements, semiconductors are usually made by doping group IV elements with group III or group V elements, so B is incorrect.
        \item Doping will increase the number of charge carriers, therefore altering the conductivity of the material, so C is correct.
        \item Electrical conductivity depends on temperature, so D is incorrect.
    \end{enumualpha}
\end{qsolution}

\begin{qsolution}{2023}{May}{Q3}{\tgElectronics \tgElectronicsSemiconductor}
    \mcqAns{B}

    When under forward bias, the n-type side of the pn junction is connected to a lower potential than the p-type side, not higher, so B is false.
\end{qsolution}

\begin{qsolution}{2023}{May}{Q4}{\tgElectronics \tgElectronicsBJT}
    \mcqAns{D}

    Assuming the BJT is npn type:
    \begin{enumualpha}
        \item When $V_{BE}<0$, the base-emitter junction is reverse biased, so no current flows through, not negative.
        \item The relationship is $I_C=\beta I_B$, not $\displaystyle I_C=\frac{I_B}{\beta}$.
        \item $I_B$ is regarded as the input current, not $I_C$.
        \item $I_E=I_B+I_C$, and $I_B$ is usually very small, so one can conclude that $I_E\approx I_C$.
    \end{enumualpha}
\end{qsolution}

\begin{qsolution}{2023}{May}{Q6}{\tgElectronics \tgElectronicsCMOS \tgElectronicsBJT}
    \mcqAns{E}

    \begin{enumualpha}
        \item BJTs are used for bipolar transistor integrated circuits, not CMOS. Incorrect.
        \item CMOS draws very low static power, but not zero. Incorrect.
        \item CMOS can be fabricated on both p-type and n-type Si substrates. Incorrect.
        \item When applied with a positive voltage at the gate, NMOS turn on, while PMOS turn off. Incorrect.
    \end{enumualpha}
\end{qsolution}

\begin{qsolution}{2023}{May}{Q7}{\tgElectronics \tgElectronicsOptoelectronics}
    \mcqAns{B}

    Photocurrent is a result of absorption of photons, which will excite an electron from the valence band to the conduction band, so B is correct.
    All other processes are unrelated to the generation of photocurrent.
\end{qsolution}

\begin{qsolution}{2023}{May}{Q8}{\tgElectronics \tgElectronicsBJT \tgElectronicsMOSFET}
    \mcqAns{E}

    MOSFETs have lower transconductance than BJTs, so E is not an advantage of MOSFETs over BJTs.
    All other options are advantages of MOSFETs over BJTs.
\end{qsolution}

\begin{qsolution}{2023}{May}{Q9}{\tgElectronics \tgElectronicsAnalogueDigital}
    \mcqAns{C}

    Both analogue and digital communication systems require modulation to transmit signals over long distances, so C is NOT exclusive to digital communication.
\end{qsolution}

\begin{qsolution}{2023}{May}{Q10}{\tgElectronics \tgElectronicsAM}
    \mcqAns{B}

    \begin{enumualpha}
        \item In SSB, LSB and carrier are suppressed, only USB is transmitted.
            Therefore, the bandwidth is halved instead of increased, so A is incorrect.
        \item The amount of noise is reduced on a narrower bandwith, so B is correct.
        \item The power dedicated to the carrier and the suppresed sideband is saved, so less power is used, so C is incorrect.
        \item There is less selective fading in SSB, so D is incorrect.
    \end{enumualpha}
\end{qsolution}

\begin{qsolution}{2023}{May}{Q11}{\tgElectronics \tgElectronicsMultiplexing}
    \mcqAns{D}

    In WDM (wavelength-division multiplexing), the varying parameter is the wavelength of the light.
    Further, we have the relationship $\displaystyle f=\frac{c}{\lambda}$, where $c$ is a constant.
    When $\lambda$ varies, $f$ (frequency) also varies, so the varying parameter is frequency.

    {\itshape
        Note: WDM can also be regarded as a type of FDM (frequency-division multiplexing).
    }
\end{qsolution}

\begin{qsolution}[100]{2023}{May}{Q12}{\tgElectronics \tgElectronicsEM}
    \mcqAns{A}

    The electric field and the magnetic field are in phase, so the phase difference is 0.
\end{qsolution}

\begin{qsolution}{2023}{May}{Q29}{\tgEE \tgEETransformer}
    \mcqAns{D}

    The AC voltage source $V_{S1}$ has a peak voltage at $\di t=\frac{1}{8}T$ (i.e., $\di \frac{2\pi}{T}=\frac{\pi}{4}$).
    $V_{S1,\mathrm{peak}}=2000\sqrt{2}$ V.
    \textbf{Note that a transformer cannot transform DC voltage (since $\di\bm{\frac{\mathrm{d}\Phi}{\mathrm{d}t}=0}$ for DC), so the DC voltage source $\bm{V_{S2}}$ does not contribute to the primary voltage.}
    By using-- \[2=\frac{N_P}{N_S}=\frac{V_P}{V_S}=\frac{I_S}{I_P},\] we have $\di V_S=\frac{1}{2}V_P=1000\sqrt{2}$ V.
    Then, since $V=IR$, we have $\di I_S=\frac{1000\sqrt{2}}{50}=20\sqrt{2}$ A.
    Using the above equation again, we have $\di I_P=\frac{1}{2}I_S=10\sqrt{2}$ A $\approx 14.14$ A.

    \textit{
        {\bfseries Note}:
        The symbols used in this solution are defined differently from the question.
        Here, $V_P$, $I_P$, $N_P$ are the voltage, current, and number of turns on the primary side, while $V_S$, $I_S$, $N_S$ are the voltage, current, and number of turns on the secondary side.
        The diagram of the question defines $I_S$ as the current on the primary side.
    }
\end{qsolution}

\begin{qsolution}{2023}{May}{Q31}{\tgEE \tgEERenewableEnergy}
    \mcqAns{E}

    Natural gas is a fossil fuel, it is consumed faster than it is replenished, so it is not a renewable energy source.
\end{qsolution}

\begin{qsolution}[20]{2023}{May}{Q33}{\tgEE \tgEEPowerSystems \tgEERMS}
    \mcqAns{E}

    The voltage waveform for one period is defined as:
    \[
        v(t)=\begin{cases}
            \frac{6t}{T}, & t\in[0, 0.5T) \\
            3, & t\in[0.5T, T)
        \end{cases}.
    \]
    Then, the RMS voltage value is given by:
    \begin{align*}
        V_{\mathrm{RMS}} &=
        \sqrt{\frac{1}{T}\int_{0}^{T} v^2(t)\d t} \\
        &= \sqrt{\frac{1}{T}\left(\int_0^{0.5T}\frac{36t^2}{T^2}\d t + \int_{0.5T}^{T} 9\d t\right)} \\
        &= \sqrt{\frac{1}{T}\left(\frac{3}{2}T+\frac{9}{2}T\right)} \\
        &= \sqrt{6}\,\mathrm{V} \approx 2.45\,\mathrm{V}
    \end{align*}
\end{qsolution}

\begin{qsolution}{2023}{May}{Q35}{\tgEE \tgEEDeltaY}
    \mcqAns{B}

    The transformation from Delta to Y is given by
    \[R_a=\frac{R_2\cdot R_3}{R_1+R_2+R_3}=\frac{4\cdot 6}{2+4+6}=2\,\Omega\]
\end{qsolution}

\end{multicols*}