\section{Number Representation}

\subsection*{Positional Number System}
$\sum_i (a_ir^i)$ \hfill Direct ($O(n^2)$)\\
$r(r(a_n + a_{n-1})+\cdots)+a_0$ \hfill Iterative ($O(n)$)\\

\subsection*{Binary Integers}
\emph{\texttt{uint}}: $\sum_{i=0}^{n-1} 2^i a_i$\\
\emph{Sign-Mag}: $(-1)^{a_{n-1}}\sum_{i=0}^{n-2} 2^i a_i$\\
\emph{1's Comp}: (if $<0$) bit-wise NOT\\
\emph{2's Comp}: (if $<0$) 1's Comp + 1 \quad $-2^{n-1} + \sum_{i=0}^{n-2} 2^i a_i$\\
MSB of 1's Comp and 2's Comp is sign bit.\\

\subsection*{Binary Integer Arithmetics}
\subsubsection*{Negation of 2's Comp}
Take 2's Comp of the 2's Comp.
\subsubsection*{Add/Sub of 2's Comp}
Add/Sub directly.\\
\emph{Overflow}: Two numbers of same sign added to get oppposite sign.
\subsubsection*{Multiplication (multiplicand $\times$ multiplier)}
\emph{$+$ve$\times+$ve}: \begin{enumerate*}[label=(\arabic*)]
    \item for each multiplier bit,
    \item if 1, shift multiplicand left, add to partial sum,
    \item if 0, do nothing,
    \item return sum.
\end{enumerate*}
\emph{Other cases}: \begin{enumerate*}[label=(\arabic*)]
    \item for each non-sign multiplier bit,
    \item if 1, shift multiplicand left, add to partial sum,
    \item if 0, do nothing,
    \item for sign bit, if 1, negate multiplicand, left shift, sign extend, add to partial sum,
    \item return sum.
\end{enumerate*}

\subsection*{Excess-$K$}
Values range: $[0-K, 2^n-1-K]$\\
$K$ is typically chosen to be $2^{n-1} - 1$.

\subsection*{Floating Point Numbers}
$\pm\text{Significand}\times2^{\pm\text{(Biased) Exponent}}$\\
\emph{Single}: 32 bits, 8 exp, 23 sig.\\
\emph{Double}: 64 bits, 11 exp, 52 sig.\\
\emph{Extended}: 80 bits, 15 exp, 112 sig.

\subsubsection*{Special Values (used only when specified)}
\emph{0}: exp = 0, sig = 0.\\
\emph{Subnormalized}: exp = 0, sig $\neq$ 0.\\
\emph{$\boldsymbol{\infty}$}: exp = all 1, sig = 0.\\
\emph{NaN}: exp = all 1, sig $\neq$ 0.

\subsubsection*{Properties}
\# of representable numbers same as int.\\
Not uniformly distributed.\\
Arithmetic laws not always hold.

\subsection*{Floating Point Arithmetics}

\subsubsection*{Add/Sub}
\begin{enumerate*}[label=(\arabic*)]
    \item Check 0.
    \item Align significand (smaller exp shift right).
    \item Add/Sub significands.
    \item Normalise.
    \item Round.
\end{enumerate*}

\subsubsection*{Multiplication/Division}
\begin{enumerate*}[label=(\arabic*)]
    \item Check 0.
    \item Multiply/Divide significands.
    \item Multiplication: add exponents, sub $K$; Division: sub exponents, add $K$.
    \item Determine sign.
    \item Normalise.
    \item Round.
\end{enumerate*}

\subsubsection*{Rounding Methods}
Round to nearest even.\\
Round towards zero.\\
Round towards $\pm\infty$.